---
title: 'PART 2: Basic Inferential Data Analysis'
author: "WendyD"
date: "9/9/2020"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(ggplot2)
```

# Load the ToothGrowth data and perform some basic exploratory data analyses

```{r, echo=TRUE}
data(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)
```

# Basic Summary of the data

```{r, echo=TRUE}
summary(ToothGrowth)
```

Since our data has 3 variables: len(tooth length), supp (supplement types), and dose (in mg/day), we could explore our data, using dosage as factor variable, and to see how the tooth length changes with different dosage accordingly to their supplement types


```{r, echo=TRUE}
ggplot(data=ToothGrowth, aes(x=factor(dose), y=len, colour=factor(dose)))+
               facet_grid(.~ToothGrowth$supp)+
               geom_boxplot()+
               xlab("Dose ")+ ylab("Tooth Length")+
               ggtitle("Tooth Length by Dosage for Every Supplement")

```

From the graph, it appreared that Orange juice outperformed ascorbic acid at dose 0.5mg/day and 1mg/day
We can also compare the effectivty of 2 supplements by calculating the mean by supplements factor
```{r,echo=TRUE}
tapply(ToothGrowth$len, ToothGrowth$supp, mean)

```

This confirm our observation, Orange juice does increase the tooth length faster than ascirbic acid


# Hypothesis Testing to compare tooth growth by supp and dose

## Analyze tooth growth by supplement

We test two hypothesis: First: there is no correlation in the tooth length for different supplements. Secondly, there is no correlation in tooth length acorss different doses of supplement.


Tooth length for those taking orange juice as supplement
```{r,echo=TRUE}
orangejuice<- subset(ToothGrowth
                     $len, ToothGrowth$supp == "OJ")
```

Tooth length for those taking ascorbic acid as supplement
```{r, echo=TRUE}
aacid<- subset(ToothGrowth$len, ToothGrowth$supp =="VC")

```

Compute t test to see if there is any significant difference in the tooth length for people taking different supplement

```{r,echo=TRUE}
t.test(orangejuice, aacid )

```


Since the p value is > than 0.05, we cannot reject the null hypothesis, and concluded that there is no significant difference in tooth length for different supplements

## Analyze tooth growth across dosages

```{r, echo=TRUE}
dose0.5<- subset(ToothGrowth$len, ToothGrowth$dose == "0.5")
dose1<- subset(ToothGrowth$len, ToothGrowth$dose == "1")
dose2<- subset(ToothGrowth$len, ToothGrowth$dose == "2")
```

Compute t test to see if there is any significant difference in the tooth length for people taking different dosesages

Dose 0.5 and dose 1mg/day

```{r, echo=TRUE}
t.test(dose0.5, dose1,paired = FALSE, alternative = "less",  var.equal = FALSE )
```

Dose 1 and dose 2
```{r,echo=TRUE}
t.test(dose1, dose2,paired = FALSE, alternative = "less",  var.equal = FALSE )

```


Dose 0.5 and Dose 2 mg/day

```{r,echo=false}
t.test(dose0.5, dose2,paired = FALSE, alternative = "less",  var.equal = FALSE )

```


Since the p value for both 3 t test are very small, we have to reject the null hypothesis and concluded that there are correlation in dosages and tooth length

# Conclusion
We concluded that there is  no significant correlation in taking orange juice or ascorbic acid in regard to tooth lengths. We analyzed the correlation between 0.5 and 1 mg/day, 0.5 and 2mg/day, and 1mg and 2mg/day. There are  strong correlation between 3 different dosages and the tooth length to one another. 
