---
title: "MachineLearning_Proj"
author: "WendyD"
date: "05/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required packages
```{r, warning=FALSE}
library(caret)
library(ggplot2)
library(tidyverse)
library(randomForest)
```

# Load Data

Read train and test set. Using excel, we saw that there is some #DIV/0! and blank value data, hence we remeove this with na.strings parameter 
```{r}
train.df<-read.csv('pml-training.csv', na.strings= c("#DIV/0!")) 
test.df <- read.csv('pml-testing.csv', na.strings = c("#DIV/0!"))
dim(train.df)
dim(test.df)
```

# Preprocessing Data

## Remove unnecessary variables which is the time frame, time window, and users

```{r}
train.df <- train.df[ , -c(1:7)]
test.df <- test.df[ , -c(1:7)]
dim(train.df)

```

## Feature Selection

Remove column has more than 95 NA and near zeri variance

```{r}
thrshold <- dim(train.df)*0.95
#remove column has 95% or more NA or ""  value
goodcolumn <- !apply(train.df, 2, function(x) sum(is.na(x))>thrshold || sum(x=="")>thrshold)
train.df <- train.df[ ,goodcolumn]

#remove near zero variance column
badcolumn <- nearZeroVar(train.df, saveMetrics = TRUE)
#having the column with near zero variance = TRUE ->> remove these column
train.df <- train.df[ , badcolumn$nzv==FALSE]

train.df$classe <- factor(train.df$classe)


#Repeat the same thing for test data set
goodcolumn2 <- !apply(test.df, 2, function(x) sum(is.na(x))>thrshold || sum(x=="")>thrshold)
test.df <- test.df[ ,goodcolumn2]
badcolumn2 <- nearZeroVar(test.df, saveMetrics = TRUE)

test.df <- test.df[ ,badcolumn2$nzv==FALSE]
testing <- test.df
testing$classe <- NA

```

## Split training set into train and validation set

```{r}
set.seed(123)
trainobject <- createDataPartition(train.df$classe, p=0.8, list=FALSE)
training <- train.df[trainobject,]
validation <- train.df[-trainobject,]
dim(training); dim(validation)
```



# Train Model
Use random forest, gbm or lda


```{r}
#Train 3 different models
#verbose(data_out = TRUE, data_in = FALSE, info = FALSE, ssl = FALSE)
mod1 <- train(classe ~ ., data=training, method="rf")  #random forest
#mod2 <- train(classe ~ ., data=training, method="gbm")
#mod3 <- train(classe ~ ., data=training, method="lda")

pred1 <- predict(mod1, validation)
#pred1.2<- predict(mod1,training)
#pred2 <- predict(mod2, validation)
#pred3 <- predict(mod3, validation)
```




```{r}
confusionMatrix(pred1, validation$classe)
#confusionMatrix(pred1.2, training$classe)

```

```{r}
pml_write_files = function(x){
n = length(x)
for(i in 1:n){
filename = paste0("problem_id_",i,".txt")
write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
}
}
x <- test.df

answer <- predict(mod1,testing)
answers
```
